---
title: "Untitled"
author: "Sam Hausner-Levine"
date: "12/23/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
    code_download: true
editor_options: 
chunk_output_type: console
---
# Table of Contents
1. Maps of Carjackings
2. Fishnet Maps of Risk Factors
3. Local Moran's I Maps
4. Correlation Scatter Plots and Table
5. Histogram of Dependent Variable
6. Regressions and Model Errors
7. Raw Errors by Race Context
8. Kernel Density vs. Risk Prediction
9. Conclusion

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

library(tidyverse)
library(sf)
library(RSocrata)
library(viridis)
library(spatstat)
library(raster)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(tidycensus)
library(ggcorrplot)
# functions
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 16,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.text.x = element_text(size = 14))
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 16,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 14)
  )
}

#load color palettes
palette5 <- c("#fef0d9","#fdcc8a","#fc8d59","#e34a33","#b30000")
paletteMap <- c("#fee5d9","#fcae91","#fb6a4a","#de2d26","#a50f15")

policeDistricts <- 
  st_read("https://data.cityofchicago.org/api/geospatial/fthy-xz3r?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = dist_num)
  
policeBeats <- 
  st_read("https://data.cityofchicago.org/api/geospatial/aerh-rz74?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = beat_num)

bothPoliceUnits <- rbind(mutate(policeDistricts, Legend = "Police Districts"), 
                         mutate(policeBeats, Legend = "Police Beats"))

crimes <-  read.socrata("https://data.cityofchicago.org/resource/w98m-zvie.json")
    
CarJacking <- 
  crimes %>% 
    filter(primary_type == "ROBBERY" & description == "VEHICULAR HIJACKING") %>%
    na.omit() %>%
    st_as_sf(coords = c("location.longitude", "location.latitude"), crs = 4326, agr = "constant")%>%
    st_transform('ESRI:102271') %>% 
    distinct()

CarJacking2020 <- read.socrata("https://data.cityofchicago.org/resource/qzdf-xmn8.json") %>%
    filter(primary_type == "ROBBERY" & description == "VEHICULAR HIJACKING") %>%
    na.omit() %>%
    st_as_sf(coords = c("location.longitude", "location.latitude"), crs = 4326, agr = "constant")%>%
    st_transform('ESRI:102271') %>% 
    distinct()


chicagoBoundary <- 
  st_read(file.path(root.dir,"/Chapter5/chicagoBoundary.geojson")) %>%
  st_transform('ESRI:102271') 

fishnet <- 
  st_make_grid(chicagoBoundary,
               cellsize = 1000, 
               square = TRUE) %>%
  .[chicagoBoundary] %>%            # <- MDH Added
  st_sf() %>%
  mutate(uniqueID = rownames(.))

Bars <- 
  read.socrata("https://data.cityofchicago.org/resource/nrmj-3kcf.json")%>%
  filter(license_description == 'Tavern')%>%
  dplyr::select(location.longitude,location.latitude)%>%
    na.omit() %>%
    st_as_sf(coords = c("location.longitude", "location.latitude"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Bars")


Liquor.Stores <- 
  read.socrata("https://data.cityofchicago.org/resource/nrmj-3kcf.json")%>% 
  filter(license_description == 'Package Goods')%>%
  dplyr::select(location.longitude,location.latitude)%>%
    na.omit() %>%
    st_as_sf(coords = c("location.longitude", "location.latitude"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Liquor Stores")


Robbery <- crimes %>% 
    filter(primary_type == "ROBBERY" & description != "VEHICULAR HIJACKING") %>%
  dplyr::select(longitude,latitude)%>%
    na.omit() %>%
    st_as_sf(coords = c("longitude", "latitude"), crs = 4326, agr = "constant")%>%
    st_transform('ESRI:102271') %>% 
  mutate(Legend = "Robberies")
  
Abandoned.Vehicles <- 
  read.socrata("https://data.cityofchicago.org/resource/3c9v-pnva.json")%>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2019") %>%
dplyr::select(Y = location.latitude, X = location.longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Abandoned Vehicles")

Out.Street.Lights <- 
  read.socrata("https://data.cityofchicago.org/resource/756k-itxx.json") 

OSL.2019 <- Out.Street.Lights %>%
  mutate(year = substr(creation_date,1,4)) %>% filter(year == "2018")%>%
dplyr::select(Y = latitude, X = longitude)%>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Out Lights")

Bus.Stops <- st_read("C:/Users/shaus/Downloads/Hospitals (1).geojson")%>%
  st_transform('ESRI:102271')%>%
    dplyr::select(geometry) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Bus Stops")

L.Stations <- st_read("C:/Users/shaus/OneDrive/Desktop/Weitzman CPLN/Fall 2021/Public Policy Analytics/gis-data/stations_metra/metra_entrances.geojson")%>%
  st_transform('ESRI:102271')%>%
    dplyr::select(geometry) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "L Stations")



Public.Schools <- st_read("C:/Users/shaus/Downloads/Chicago Public Schools - School Locations SY1819.geojson")%>%
  st_transform('ESRI:102271')%>%
    dplyr::select(geometry) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Public Schools")



neighborhoods <- 
  st_read("https://raw.githubusercontent.com/blackmad/neighborhoods/master/chicago.geojson") %>%
  st_transform(st_crs(fishnet)) 
```

# Maps of Carjackings
## Point Form and Density

The maps in this section show the location and density of reported car jackings in Chicago in 2019. The maps show a major density spot to the west of the city and another spot to the south. Because of the relative infrequency of car jackings in comparison to other crimes, note that there are large sections of the city in the northwest and northeast that have no reported car jackings at all.

```{r Point Form and Density Maps, message=FALSE, warning=FALSE, include=TRUE, results='markup'}
grid.arrange(ncol=2,
ggplot() + 
  geom_sf(data = chicagoBoundary) +
  geom_sf(data = CarJacking, colour="red", size=0.1, show.legend = "point") +
  labs(title= "Figure 1.1: Car Jackings, Chicago - 2019") +
  mapTheme(),

ggplot() + 
  geom_sf(data = chicagoBoundary, fill = "grey40") +
  stat_density2d(data = data.frame(st_coordinates(CarJacking)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_viridis() +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Figure 1.2: Density of Car Jacking") +
  mapTheme() + theme(legend.position = "none"))
```

## Fishnet Map

The map below shows the occurrence of carjackings grouped into a fishnet measuring 1000 feet across. Because of the lower count of crimes, I expanded the size of the fishnet to have more cells with non-zero totals.
```{r Fishnet Maps, message=FALSE, warning=FALSE, include=TRUE, results='markup'}



crime_net <- 
  dplyr::select(CarJacking) %>% 
  mutate(countCarJacking = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countCarJacking = replace_na(countCarJacking, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

ggplot() +
  geom_sf(data = crime_net, aes(fill = countCarJacking), color = NA) +
  scale_fill_viridis() +
  labs(title = "Figure 1.3: Count of Burglaires for the fishnet") +
  mapTheme()

```

# Fishnet Maps of Risk Factors
## Risk Factors Fishnet

I chose eight risk factors for the model mapped below: 

-Bars
-Liquor Stores
-Robberies(non-vehicular)
-Abandoned Vehicles
-Bus Stops
-L stations
-Out street lights
-Public Schools

Five of the factors(Bars, Liquor Stores, Robberies, Out Street Lights and Abandoned Vehicles) I chose anticipating a positive correlation with carjackings because of general negative impact on public safety.

The others were chosen assuming they have an effect on traffic patterns in the area, although without any assumption of whether the correlation would be positive or negative.

A cursory comparison of maps shows that map of distance to the nearest bus stop overlaps significantly with the density of car jackings, meaning that in areas where there are more bus stops there are also tend to be more carjackings. 

``` {r Risk Factor Fishnet , message=FALSE, warning=FALSE, include=TRUE, results='markup'}
vars_net <- rbind(Bars, Liquor.Stores, Robbery, Abandoned.Vehicles, Bus.Stops, L.Stations, OSL.2019, Public.Schools) %>%
  st_join(., fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
  full_join(fishnet, by = "uniqueID") %>%
  spread(Legend, count, fill=0) %>%
  st_sf() %>%
  dplyr::select(-`<NA>`) %>%
  na.omit() %>%
  ungroup()

st_c    <- st_coordinates
st_coid <- st_centroid

vars_net <- vars_net %>%
    mutate(
           Liquor.Stores.nn = nn_function(st_c(st_coid(vars_net)), 
                                           st_c(Liquor.Stores),
                                           k = 5),
           Robbery.nn = nn_function(st_c(st_coid(vars_net)), 
                                           st_c(Robbery),
                                           k = 5),
           Abandoned.Vehicles.nn = nn_function(st_c(st_coid(vars_net)), 
                                           st_c(Abandoned.Vehicles),
                                           k = 5),
           Bus.Stops.nn = nn_function(st_c(st_coid(vars_net)), 
                                           st_c(Bus.Stops),
                                           k = 5),
           L.Stations.nn = nn_function(st_c(st_coid(vars_net)), 
                                           st_c(L.Stations),
                                           k = 2),
           OSL.nn = nn_function(st_c(st_coid(vars_net)), 
                                           st_c(OSL.2019),
                                           k = 5),
           Public.Schools.nn = nn_function(st_c(st_coid(vars_net)), 
                                           st_c(Public.Schools),
                                           k = 2))

vars_net.long <- vars_net %>%
  dplyr::select("Abandoned Vehicles","Bus Stops", "Robberies","L Stations","Public Schools", "Out Lights", "Bars")  %>%
    gather(Variable, value, -geometry)

vars <- unique(vars_net.long$Variable)
mapList <- list()

for(i in vars){
  mapList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(vars_net.long, Variable == i), aes(fill=value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme()}

do.call(grid.arrange,c(mapList, ncol = 3, top = "Figure 2.1: Risk Factors by Fishnet"))
```

## Nearest Neighborbor Risk Factor Fishnet

```{r Nearest Neighbor Risk Factor Fishnet , message=FALSE, warning=FALSE, include=TRUE, results='markup'}
vars_net.long.nn <- 
  dplyr::select(vars_net, ends_with(".nn")) %>%
    gather(Variable, value, -geometry)

vars.nn <- unique(vars_net.long.nn$Variable)
mapList.nn <- list()

for(i in vars.nn){
  mapList.nn[[i]] <- 
    ggplot() +
      geom_sf(data = filter(vars_net.long.nn, Variable == i), aes(fill=value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme()}

do.call(grid.arrange,c(mapList.nn, ncol = 3, top = "Figure 2.2: Nearest Neighbor Risk Factors by Fishnet"))
```

# Local Moran's I Maps

The maps below show the local moran's I for carjackings in the city, or the randomness of local clustering. As the density map in section 2 showed there are significant "hot spots" immediately to the south and west of Downtown with high Moran's I scores. In these area, the number of carjackings in each cell is positively affected by the counts in the adjacent cells. The hotspot to the west of downtown appears to have a significantly higher local effect than the hotspot to the south. 


```{r Local MoranI , message=FALSE, warning=FALSE, include=TRUE, results='markup'}
 
final_net <-
  left_join(crime_net, st_drop_geometry(vars_net), by="uniqueID") 


final_net <-
  st_centroid(final_net) %>%
    st_join(dplyr::select(neighborhoods, name), by = "uniqueID") %>%
    st_join(dplyr::select(policeDistricts, District), by = "uniqueID") %>%
      st_drop_geometry() %>%
      left_join(dplyr::select(final_net, geometry, uniqueID)) %>%
      st_sf() %>%
  na.omit()

final_net.nb <- poly2nb(as_Spatial(final_net), queen=TRUE)

final_net.weights <- nb2listw(final_net.nb, style="W", zero.policy=TRUE)


local_morans <- localmoran(final_net$countCarJacking, final_net.weights, zero.policy=TRUE) %>% 
  as.data.frame()

final_net.localMorans <- 
  cbind(local_morans, as.data.frame(final_net)) %>% 
  st_sf() %>%
  dplyr::select(CarJacking_Count = countCarJacking, 
                Local_Morans_I = Ii, 
                P_Value = `Pr(z != E(Ii))`) %>%
  mutate(Significant_Hotspots = ifelse(P_Value <= 0.001, 1, 0)) %>%
  gather(Variable, Value, -geometry)
  
final_net <- final_net %>% 
  mutate(Carjacking.isSig = 
           ifelse(local_morans[,5] <= 0.001, 1, 0)) %>%
  mutate(Carjacking.isSig.dist = 
           nn_function(st_c(st_coid(final_net)),
                       st_c(st_coid(filter(final_net, 
                                           Carjacking.isSig == 1))), 
                       k = 1))

vars <- unique(final_net.localMorans$Variable)
varList <- list()

for(i in vars){
  varList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(final_net.localMorans, Variable == i), 
              aes(fill = Value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme() + theme(legend.position="bottom")}

do.call(grid.arrange,c(varList, ncol = 4, top = "Figure 3.1 - 3.4: Local Morans I statistics, Car Jackings"))
```

# Correlation Scatter Plots and Table

Of the independent variables, the count of local robberies and a location in an area of significant carjackings were unsurprisingly the ones with the strongest positive correlations. Less expected, the count of out street lights and public school buildings also had moderate correlations(.27 and .25 respectively) with car jackings. Converserly, the distance to these features had the strongest negative correlation coefficients.

As speculated before, the distance to the nearest bus stop also had a strong negative correlation with the count of carjackings in the area(-.27), although the presence of a bus station in an area itself did no have a significant correlation. 
```{r Variable Correlation, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=TRUE, results='markup'}

correlation.long <-
  st_drop_geometry(final_net) %>%
    dplyr::select(-uniqueID, -cvID) %>%
    gather(Variable, Value, -countCarJacking)

correlation.cor <-
  correlation.long[correlation.long$Variable != "name",] %>%
    group_by(Variable) %>%
    summarize(correlation = cor(as.numeric(Value), countCarJacking, use = "complete.obs"))
    
ggplot(correlation.long, aes(Value, countCarJacking)) +
  geom_point(size = 0.1) +
  geom_text(data = correlation.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "black") +
  facet_wrap(~Variable, ncol = 3, scales = "free") +
  labs(title = "Figure 4.1: Car Jacking count as a function of risk factors") +
  plotTheme()

correlation.cor <- correlation.cor[order(correlation.cor$correlation, decreasing = TRUE),]

kable(correlation.cor) %>%
 kable_styling() %>%
 footnote(general_title = "\n",
       general = "Figure 4.2: Independent Variables by Correlation")

```

# Histogram of Dependent Variable

This histogram of the car jacking count by fishnet cell shows the vast majority of cells without reported incidents and a downward curve of smaller cell counts as the number of incidents increases.

``` {r Histograms, message=FALSE, warning=FALSE, include=TRUE, results='markup'}

ggplot(data = final_net, aes(x = countCarJacking))+
  geom_histogram()+
  labs(title = "Figure 5.1: Histogram of Dependent Variables")+
  plotTheme()
```

# Regressions and Model Errors

Figure 5.1 shows the distribution of MAE for all four regressions. For the K-fold regression, adding spatial factors appears to have slightly improved the model, although for the LOGO regression the impact appears marginal. 

Figure 5.2 compares the mean MAE for all four regressions. The mean MAE for the K-fold is low(.13) in comparison for LOGO (.23), indicating significantly higher accuracy for the first model. Adding spatial factors did not significantly improve either model. 
```{r Regressions and Cross Validation, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=FALSE, results='hide'}



reg.vars <- c("Liquor.Stores.nn", "Robbery.nn", "Abandoned.Vehicles.nn", "Bus.Stops.nn", "L.Stations.nn", "OSL.nn", "Public.Schools.nn")
   
reg.ss.vars <- c("Liquor.Stores.nn", "Robbery.nn", "Abandoned.Vehicles.nn", "Bus.Stops.nn", "L.Stations.nn", "OSL.nn", "Public.Schools.nn","Carjacking.isSig","Carjacking.isSig.dist")


# CV function
crossValidate <- function(dataset, id, dependentVariable, indVariables) {
  
  allPredictions <- data.frame()
  cvID_list <- unique(dataset[[id]])
  
  for (i in cvID_list) {
    
    thisFold <- i
    cat("This hold out fold is", thisFold, "\n")
    
    fold.train <- filter(dataset, dataset[[id]] != thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, indVariables, dependentVariable)
    fold.test  <- filter(dataset, dataset[[id]] == thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, indVariables, dependentVariable)
    
    regression <- glm(paste0(dependentVariable,"~."), family = "poisson", 
    data = fold.train %>% dplyr::select(-geometry, -id))
    
    thisPrediction <- 
      mutate(fold.test, Prediction = predict(regression, fold.test, type = "response"))
    
    allPredictions <-
      rbind(allPredictions, thisPrediction)
    
  }
  return(st_sf(allPredictions))
}

reg.cv <- crossValidate(
  dataset = final_net,
  id = "cvID",
  dependentVariable = "countCarJacking",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = cvID, countCarJacking, Prediction, geometry)

reg.ss.cv <- crossValidate(
  dataset = final_net,
  id = "cvID",
  dependentVariable = "countCarJacking",
  indVariables = reg.ss.vars) %>%
    dplyr::select(cvID = cvID, countCarJacking, Prediction, geometry)
  
reg.spatialCV <- crossValidate(
  dataset = final_net,
  id = "name",
  dependentVariable = "countCarJacking",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = name, countCarJacking, Prediction, geometry)

reg.ss.spatialCV <- crossValidate(
  dataset = final_net,
  id = "name",
  dependentVariable = "countCarJacking",
  indVariables = reg.ss.vars) %>%
    dplyr::select(cvID = name, countCarJacking, Prediction, geometry)



reg.summary <- 
  rbind(
    mutate(reg.cv,           Error = Prediction - countCarJacking,
                             Regression = "Random k-fold CV: Just Risk Factors"),
                             
    mutate(reg.ss.cv,        Error = Prediction - countCarJacking,
                             Regression = "Random k-fold CV: Spatial Process"),
    
    mutate(reg.spatialCV,    Error = Prediction - countCarJacking,
                             Regression = "Spatial LOGO-CV: Just Risk Factors"),
                             
    mutate(reg.ss.spatialCV, Error = Prediction - countCarJacking,
                             Regression = "Spatial LOGO-CV: Spatial Process")) %>%
    st_sf() 

error_by_reg_and_fold <- 
  reg.summary %>%
    group_by(Regression, cvID) %>% 
    summarize(Mean_Error = mean(Prediction - countCarJacking, na.rm = T),
              MAE = mean(abs(Mean_Error), na.rm = T),
              SD_MAE = mean(abs(Mean_Error), na.rm = T)) %>%
  ungroup()
```

## Distribution of MAE
```{r MAE and Goodness of Fit Hist, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=TRUE, results='markup'}
error_by_reg_and_fold %>%
  ggplot(aes(MAE)) + 
    geom_histogram(bins = 30, colour="black", fill = "#FDE725FF") +
    facet_wrap(~Regression) +  
    geom_vline(xintercept = 0) + scale_x_continuous(breaks = seq(0, 8, by = 1)) + 
    labs(title="Figure 5.1: Distribution of MAE", subtitle = "k-fold cross validation vs. LOGO-CV",
         x="Mean Absolute Error", y="Count") +
    plotTheme()
```

## MAE by Regression
```{r MAE and Goodness of Fit Table, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, include=TRUE, results='markup'}
st_drop_geometry(error_by_reg_and_fold) %>%
  group_by(Regression) %>% 
    summarize(Mean_MAE = round(mean(MAE), 2),
              SD_MAE = round(sd(MAE), 2)) %>%
  kable(caption = "Figure 5.2: MAE by Regression") %>%
    kable_styling("striped", full_width = F) %>%
    row_spec(2, color = "black", background = "#FDE725FF") %>%
    row_spec(4, color = "black", background = "#FDE725FF") 


```

## Raw Errors by Race Context

When the regression was grouped by race context, MAE changed somewhat. With just risk factors, the regression slighlty underpredicted for non-white majority areas, and slightly over predicted for white majority areas. When spatial factors were introduced, this relationship was inverted, with predictions in non-white areas slightly exceeding counts. Generally, the model appears to have low levels of bias. 


```{r Raw Errors by Race Context, message=FALSE, warning=FALSE, include=TRUE, results='markup'}

tracts19 <- 
  get_acs(geography = "tract", variables = c("B01001_001E","B01001A_001E"), 
          year = 2018, state=17, county=031, geometry=T) %>%
  st_transform('ESRI:102271')  %>% 
  dplyr::select(variable, estimate, GEOID) %>%
  spread(variable, estimate) %>%
  rename(TotalPop = B01001_001,
         NumberWhites = B01001A_001) %>%
  mutate(percentWhite = NumberWhites / TotalPop,
         raceContext = ifelse(percentWhite > .5, "Majority_White", "Majority_Non_White")) %>%
  .[neighborhoods,]

reg.summary %>% 
  filter(str_detect(Regression, "LOGO")) %>%
    st_centroid() %>%
    st_join(tracts19) %>%
    na.omit() %>%
      st_drop_geometry() %>%
      group_by(Regression, raceContext) %>%
      summarize(mean.Error = mean(Error, na.rm = T)) %>%
      spread(raceContext, mean.Error) %>%
      kable(caption = "Mean Error by neighborhood racial context") %>%
        kable_styling("striped", full_width = F)  
```

# Kernel Density vs. Risk Prediction
The maps below show kernel density and local risk prediction from the 2019 model. In figure 6.2, car jackings from 2020 are overlaid on top of the kernel density and local risk prediction maps to compare predictive ability. As shown in figure 6.3, the two highest risk areas of the risk prediction map captured more of the 2020 car jackings than the for kernel density map, showing that the predictive value of the model is stronger than following crime density alone. 
```{r Kernel Density and Risk Prediction, message=FALSE, warning=FALSE, include=TRUE, results='markup'}

CJ_ppp <- as.ppp(st_coordinates(CarJacking), W = st_bbox(final_net))
CJ_KD.1000 <- spatstat.core::density.ppp(CJ_ppp, 1000)



as.data.frame(CJ_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) %>%
   ggplot() +
     geom_sf(aes(fill=value)) +
     scale_fill_viridis(name = "Density") +
     labs(title = "Figure 6.1: Kernel density of 2019 Carjackings") +
     mapTheme()




CJ_KDE_sf <- as.data.frame(CJ_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) %>%
  mutate(label = "Kernel Density",
         Risk_Category = ntile(value, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 70 & Risk_Category <= 89 ~ "70% to 89%",
           Risk_Category >= 50 & Risk_Category <= 69 ~ "50% to 69%",
           Risk_Category >= 30 & Risk_Category <= 49 ~ "30% to 49%",
           Risk_Category >= 1 & Risk_Category  <= 29 ~ "1% to 29%")) %>%
  cbind(
    aggregate(
      dplyr::select(CarJacking2020) %>% mutate(CJCount = 1), ., sum) %>%
    mutate(CJCount = replace_na(CJCount, 0))) %>%
  dplyr::select(label, Risk_Category, CJCount)

CJ_risk_sf <-
  filter(reg.summary, Regression == "Random k-fold CV: Spatial Process") %>%
  mutate(label = "Risk Predictions",
         Risk_Category = ntile(Prediction, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 70 & Risk_Category <= 89 ~ "70% to 89%",
           Risk_Category >= 50 & Risk_Category <= 69 ~ "50% to 69%",
           Risk_Category >= 30 & Risk_Category <= 49 ~ "30% to 49%",
           Risk_Category >= 1 & Risk_Category <= 29 ~ "1% to 29%")) %>%
  cbind(
    aggregate(
      dplyr::select(CarJacking2020) %>% mutate(CJCount = 1), ., sum) %>%
      mutate(CJCount = replace_na(CJCount, 0))) %>%
  dplyr::select(label,Risk_Category, CJCount)

#Making comparisons for 2017 vs 2018

rbind(CJ_KDE_sf, CJ_risk_sf) %>%
  na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category, -geometry) %>%
  ggplot() +
    geom_sf(aes(fill = Risk_Category), colour = NA) +
    geom_sf(data = sample_n(CarJacking2020, 3000,replace = TRUE), size = .5, colour = "black") +
    facet_wrap(~label, ) +
    scale_fill_viridis(discrete = TRUE) +
    labs(title="Figure 6.2: Comparison of Kernel Density and Risk Predictions",
         subtitle="2019 car jacking risk predictions; 2020 car jacking overdoses") +
    mapTheme()

##The bar plot making this comparison

rbind(CJ_KDE_sf, CJ_risk_sf) %>%
  st_set_geometry(NULL) %>% na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category) %>%
  group_by(label, Risk_Category) %>%
  summarize(countCJs = sum(Value)) %>%
  ungroup() %>%
  group_by(label) %>%
  mutate(Rate_of_test_set_crimes = countCJs / sum(countCJs)) %>%
    ggplot(aes(Risk_Category,Rate_of_test_set_crimes)) +
      geom_bar(aes(fill=label), position="dodge", stat="identity") +
      scale_fill_viridis(discrete = TRUE) +
      labs(title = "Figure 6.3: Risk prediction vs. Kernel density, 2020 Car Jackings") +
      plotTheme() + theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

```


# Conclusion

  I would recommend using the algorithim produced in this exercise given its strong predicitive ability and low margins of error. As shown in the final section(figures 6.2 and 6.3), using this algorithim was more accurate than looking at crime density from the previous year alone, particularly in high risk areas. If the Chicago PD were to implement this algorithm to deploy limited resources, they would have better results than deploying units simply based on the density of incidents from the year before.

 Importantly, the model showed strong generalizability across different neighborhoods. When cross validated using the k-fold method, the margins of error were fairly low, indicating good generalizability. The model also showed relatively low levels of racial bias. Without including spatial factors, the regression led to under predicting incidents in non-white majority neighborhoods, and margins of error in both categories were low. When spatial risk factors were included, the model began to over predict for non-white majority neighborhoods, although only slightly. This may indicate that car jacking "hot spots" overlap with high concentrations of non-white residents. While the algorithm itself does not appear to have biased results, it would be important to further understand the relationship between racial context and the model's spatial risk factors before implementing the program. 
